---
title: "WUP_04"
author: "Monserrat Mendez"
date: "spring 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

GSScsv <- read.csv(file= "GSScsv.csv")

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
basicCN <- ggplot(GSScsv,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```
This graph shows that there's some significant statistical differences among this collection of numerical explained by categorical variables. as all the different school levels seem to be in different to be in different areas some meet. however two of them stick out the most that being "Graduate" and "High school" in which graduates has a lighter leve while high school as one of if not the lowest 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSScsv)
```
Based on the data, it appears that graduates have the highest median value compared to all other categories. Additionally, the third quarterly values of high school, students, and light high school students are relatively close and smaller than all the other values. It is worth noting that light high school students have the overall lowest median and mean values.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv)
summary.aov(model1)
```
```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 18.7, 2, 2))
plot(TukeyHSD(model1),las=2)
```
The graph shows that high school and graduate school are significantly different.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


The graph shows significant differences in numerical values for different education levels. Graduates have the highest median value, while light high school students have the lowest median and mean values. The graph also highlights a substantial contrast between high school and graduate school. Overall, this information can be useful for informed decision-making and further analysis.